---
title: "Homework 2: Evaluating Randomized Experiments"
author: "Neeraj Sharma"
date: "05/02/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r include = FALSE}
library(tidyverse)
library(knitr)
library(magrittr)
library(ggthemes)
```

In 1 and 3, assume randomization worked. For the rest, then you can't assume randomization worked. 
How to estimate if it's random. Multiple linear regression isn't awesome for prediction but we should use it here for randomness assessment. 

## Question 1: Estimate the treatment effect and the associated standard error on the raw data given to you, assuming there are no mistakes in or problems with the data. 

By definition, the Average Treatment Effect is defined to be $\frac{1}{N}\sum_i y_1(i) - y_0(i)$ where $y_1(i)$ and $y_0(i)$ are the values of the outcome variable (in this case hours exercised) in each treatment scenario. Unfortunately, it is often impractical to utilize this straightforward approach. This formula presumes that one can quantify the outcome of a given individual when treatment is both given and withheld. However, because each individual can only be slotted into one category, this approach cannot be perfectly achieved. 

Thus, a random experiment with a control and treatment group can be conducted to smooth out differences among populations in order to isolate the treatment effect specifically. With large enough sample groups, the difference between the mean of the outcome of the treatment group and mean of the outcome of the control group yields the Average Treatment Effect. 

Here are the relevant summary statistics of the data set without any modification. 

```{r message=FALSE}
# Read in the raw data. 
raw <-  read_csv("problem_set2.csv")

# Create dataframe with summary statistics for hours
unclean_data <- raw %>%
  group_by(treatment) %>%
  summarize(count = n(), mean(hours), sd(hours), se = sd(hours)/sqrt(n()))

# Kable print unclean_data
unclean_data %>%
  knitr::kable(col.names = c("Treatment", "Count", "Mean", "St Dev", "St Err"), 
        caption = "Summary Statistics of Hours Exercised across Sample Groups with No Data Cleaning")
```

The mean hours exercised for individuals in the treatment group is `r unclean_data %>% extract(2, 3)` and the mean hours exercised for individuals in the non-treatment group is `r unclean_data %>% extract(1, 3)`. Assuming randomness was properly implimented in this study, the difference between these two numbers will be the Average Treatment Effect of the treatment based on the analysis I provide above. Thus, the Average Treatment Effect is `r unclean_data %>% extract(2, 3) - unclean_data %>% extract(1, 3)` with a standard error of `r round(unclean_data %>% extract(2, 5) - unclean_data %>% extract(1, 5), digits = 4)`.

## Question 2: It’s always a good idea to check a data set for errors. Clean this data set as you think appropriate. As an answer to this question, note all the kinds of changes you made to the data, a few words explaining your reason for the change, and which observations you changed (noting the observation number included as a variable in the data set for identification purposes).  If it is totally obvious which observations you changed and there are a large number in the category (e.g. if you decided to drop all White study participants), you can just note what you did for that change (“I dropped all white participants”) and explain why. 

```{r}
tibble(
  Variable = append(names(raw), "changed?"),
  `Description of Modification` = c(
    "No change",
    "Hours over 60 are minutes; reformatted to be hours.",
    "No change.",
    "No change.",
    "Uniformly store data as 0/1, not 0/1/male/female.",
    "-99 means missing age data; recoded to be NA.",
    "BMIs of less than 1 are omitted.",
    "No change.",
    "Race/Ethnicity BLACK capitalization fixed.",
    "`changed?` counts changes that occured in cleaning an observation."
  )
) %>%
  knitr::kable()
```

```{r echo = TRUE}
# Clean the data to impose uniformity upon the variable encoding.  
clean_data <- raw %>%
  mutate(`changed?` = seq(1, 1000) * 0) %>%
  # Fixing messed up hours readings. They start at 60 and upwards and that's 1 hr so I
  # fix based on that.
  mutate(`changed?` = if_else(hours >= 60, `changed?` + 1, `changed?`),
         hours = if_else(hours >= 60, hours / 60, hours)) %>%
  mutate(`changed?` = if_else(community_center %in% c("WOODLAWN",
                                                      "hyde park",
                                                      "woodlawn",
                                                      "Hyd Park",
                                                      "HYDE PARK",
                                                      "Hyde_Park",
                                                      "HYDE_PARK",
                                                      "hyde_park"), 
                              `changed?` + 1, `changed?`),
         community_center = if_else(community_center %in% c("WOODLAWN", "woodlawn"),
                                    "Woodlawn", community_center),
         community_center = if_else(community_center %in% c("hyde park", 
                                                            "Hyd Park", 
                                                            "HYDE PARK", 
                                                            "HYDE_PARK", 
                                                            "hyde_park",
                                                            "Hyde_Park"), 
                                    "Hyde Park", community_center)) %>%
  # Recoding female variable to be factor categorical variable from 0/1/male/female.
  mutate(`changed?` = if_else(female %in% c("female", "male"), 
                              `changed?` + 1, `changed?`),
         female = as.double(if_else(female == "female", 
                                    "1", if_else(female == "male", "0", female)))) %>%
  # -99 is missing age data so I reincode it at missing age data. 
  # https://cran.r-project.org/web/packages/naniar/vignettes/replace-with-na.html
  mutate(`changed?` = if_else(age == -99, `changed?` + 1, `changed?`), 
         age = na_if(age,-99)) %>%
  # Currently I have removed improper values, but I could also justify multiplying by 10.
  mutate(`changed?` = if_else(bmi < 1, `changed?` + 1, `changed?`),
         bmi = ifelse(bmi < 1, NA, bmi)) %>%
  # Fix all BLACK observations to normal capitalization structure.
  mutate(`changed?` = if_else(is.na(race_ethnicity), 
                              `changed?`, if_else(race_ethnicity == "BLACK", 
                                                  `changed?` + 1, `changed?`)),
    race_ethnicity = if_else(is.na(race_ethnicity), 
                             race_ethnicity, str_to_sentence(race_ethnicity))) # %>%
# mutate(race_ethnicity = factor(race_ethnicity, c("Black", "Hispanic", "White"))) %>%
# mutate(treatment = factor(treatment, labels = c("No Treatment ($0)", "Treatment ($10)"))) %>%
# mutate(community_center = factor(community_center, labels = c("Woodlawn", "Hyde Park"))) %>%
# mutate(education = factor(education, levels = c("less than high school",
#                                                 "high school",
#                                                 "higher degree")))
```

## Question 3: With your cleaned data set, re-estimate the treatment effect and estimated standard error, assuming the randomization worked fine. 

It is clear that there are numerous outliers in terms of BMI data. Given that a BMI of less than 18.5 is underweight, having a BMI of ~1 is underweight to the point of impossibility. Thus, I believe these data points were improperly encoded and given the distribution, they appear to have simply misplaced the decimal point two places to the left.

## Question 4:Evaluate whether the randomization appears legitimate. If you don’t think the randomization is legitimate, what is your evidence? (Hint: something went wrong.)

If there was perfect random assignment, you could randomly pluck 100 people, then pluck 100 more, their distributions should be the same of age BMI Community etc. Systematic differences imply  that it's not random. Talk about mean median and stuff but you don't need to do a t-test. 

How to tell if you aren't random. Clues that it is the case is done through a balance table. Compare means of each observable metric between treatment and control groups. You do a t-test to compare them. THIS IS NOT GREAT, but still do it to confirm you are balanced on your observables. 

```{r}
# Checking randomization

# Balance Table
clean_data %>% 
  group_by(treatment) %>%
  summarize_all(mean) %>%
  kable()
```

```{r}
# Distribution of different races
clean_data %>%
  group_by(treatment) %>%
  count(race_ethnicity) %>%
  mutate(race_ethnicity = as.character(race_ethnicity)) %>%
  pivot_wider(treatment, names_from = race_ethnicity, values_from = n, values_fn = list(n = list)) %>%
  unnest() %>%
  kable(caption = "Distribution of Race/Ethnicities Sampled in Each  Group")
```

```{r}
randomness <- clean_data %>%
  mutate(treatment = factor(treatment, labels = c("No Treatment ($0)", "Treatment ($10)")))

randomness %>%
  drop_na(bmi) %>%
ggplot(mapping = aes(x = bmi)) +
  geom_histogram(fill = "#3a89cf", color = "black", bins = 15) +
  labs(x = "Body Mass Index", y = "Count", title = "Histogram of distribution of BMI between treatment groups") +
  theme_economist() +
  facet_wrap(~ treatment)

randomness %>%
  drop_na(bmi) %>%
  group_by(treatment) %>%
  summarize(count = n(), mean(bmi), sd(bmi), se = sd(bmi)/sqrt(n())) %>%
  kable(col.names = c("Treatment", "Count", "Mean", "St Dev", "St Err"), caption = "Summary Statistics of BMI across Sample Groups")


randomness %>%
  drop_na(age) %>%
ggplot(mapping = aes(x = age)) +
  geom_histogram(fill = "#3a89cf", color = "black", bins = 15) +
  labs(x = "Age", y = "Count", title = "Histogram of distribution of Age between treatment groups") +
  theme_economist() +
  facet_wrap(~ treatment)

randomness %>%
ggplot(mapping = aes(x = hours)) +
  geom_histogram(fill = "#3a89cf", color = "black", bins = 15) +
  labs(x = "Hours", y = "Count", title = "Histogram of distribution of Hours between treatment groups") +
  theme_economist() +
  facet_wrap(~ treatment)
```


## Question 5: Offer your best hypothesis/hypotheses as to what went wrong with the randomization?  What evidence do you have to support your hypothesis(es)?  For each of these hypotheses, describe your best strategy for estimating a plausible treatment effect, in spite of the bad randomization. (But don’t actually estimate that treatment effect.)


# Multiple Linear Regression Example
fit <- lm(y ~ x1 + x2 + x3, data=mydata)
summary(fit) # show results
anova function can be used to compare different linear models


```{r}
treatment_bmi <- glm(treatment ~ bmi, data = clean_data, family = binomial)
summary(treatment_bmi)

test <- broom::augment(treatment_bmi, type.predict = "response") %>%
  mutate(.pred = as.numeric(.fitted > .5))

mean(test$treatment != test$.pred, na.rm = TRUE)


ggplot(data = clean_data %>% filter(age > 0), mapping = aes(x = age, y = treatment)) + 
  geom_point() +
  geom_smooth(method = "lm", method.args = list(family = "binomial"), se = FALSE) +
  labs(title = "treatment/age")

clean_data %>% 
  filter(bmi > 10, age > 0) %>%
  mutate(female = factor(female, levels = c(0, 1), labels = c("Male", "Female"))) %>%
ggplot(mapping = aes(x = bmi, y = treatment, color = female)) + 
  geom_point() +
  geom_smooth(method = "lm", method.args = list(family = "binomial"), se = FALSE) +
  scale_color_discrete(name = "Sex") +
#  scale_y_discrete(breaks = c(0, 1), limits = c(0, 1)) +
  labs(title = "treatment/bmi over sex obs")

ggplot(data = clean_data %>% filter(bmi > 10, age > 0), mapping = aes(x = hours, y = treatment, color = female)) + 
  geom_point() +
  geom_smooth(method = "lm", method.args = list(family = "binomial"), se = FALSE) +
  labs(title = "sex/ treatment")
```

```{r}
clean_data %>%
  ggplot(mapping = aes(x = age, y = hours)) +
  geom_point()
```
